\documentclass[10pt, xcolor=table]{beamer}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\newcommand*\themecol{\usebeamercolor[fg]{structure}}

\setbeamertemplate{navigation symbols}{}
 \setbeamertemplate{footline}[frame number]

\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}

\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}
\tikzstyle{prob} = [rectangle, minimum width=3cm, text width = 4.5cm, minimum height=1cm, text centered, draw=black, fill= blue!20]
\tikzstyle{stat} = [rectangle, minimum width=3cm,  text width = 4.5cm, minimum height=1cm, text centered, draw=black, fill= red!20]
\tikzstyle{arrow} = [thick,->,>=stealth]


\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}

\title{STAT 111\\
{\small Recitation 3}}

\author{Mo Huang}
\institute{Email: mohuang@wharton.upenn.edu \\
\vspace{0.25cm}
Office Hours: Wednesdays 3:00 - 4:00 pm, JMHH F96\\
\vspace{0.25cm}
Slides: \url{github.com/mohuangx/STAT111-Spring2019} }


\date{February 8, 2019}


\begin{document}

\begin{frame}
\titlepage
\end{frame}



\begin{frame}{Random Variables: Questions}
\begin{itemize} 
\item[Q1:] Let $X$ be a random variable with the below distribution. Find the mean and the variance of $X$ using formula (1) and then formula (2).
\begin{table}
\begin{tabular}{|c|c|c|c|c|}
\hline
\rowcolor{blue!10} $x$ & -3 & -1 & 4&5  \\ \hline
$P(X = x)$ & 0.1 & 0.3 & 0.4 &0.2 \\
\hline
\end{tabular}
\caption{Probability distribution of $X$.} 
\end{table}
\item<2->[A1:]  {\color{red} $\mu = -3\times 0.1 -1\times 0.3 + 4\times 0.4 + 5\times 0.2 =2.$
\begin{align*}
\sigma^2&= (-3-2)^2(0.1) + (-1-2)^2(0.3) + (4-2)^2(0.4) + (5-2)^2(0.2) \\
&= 8.6.\\
  \\
\sigma^2 &= (-3)^2(0.1) + (-1)^2(0.3) + 4^2(0.4) + 5^2(0.2) - 2^2\\
&=8.6.
\end{align*}}
\end{itemize} 

\end{frame}

\begin{frame}{Random Variables: Questions}
\begin{itemize} 
\item[Q2:] Let $X$ be a random variable with the below distribution. Write the probability distribution of $Y = 2X$ in tableau form.
\bigskip
\begin{table}
\begin{tabular}{|c|c|c|c|c|}
\hline
\rowcolor{blue!10} $x$ & -3 & -1 & 4&5  \\ \hline
$P(X = x)$ & 0.1 & 0.3 & 0.4 &0.2 \\
\hline
\end{tabular}
\caption{Probability distribution of $X$.} 
\end{table}
\item<2->[A2:]  {\color{red} Probability distribution of $Y$:}
\bigskip
\begin{table}[H]
\begin{tabular}{|c|c|c|c|c|}
\hline
\rowcolor{blue!10} $y$ & -6 & -2 & 8 & 10  \\ \hline
$P(Y = y)$ & 0.1 & 0.3 & 0.4 &0.2 \\
\hline
\end{tabular}
\caption{Probability distribution of $Y$.} 
\end{table}
\end{itemize} 

\end{frame}



\begin{frame}{Questions}
\begin{itemize}
\setlength{\itemsep}{15pt}
\item[Q4:] Suppose we have a company producing a medicine. Each day the \emph{mean} amount of medicine produced is 500 mg and the \emph{variance} is 900 mg$^2$. Assume the amount produced each day is independent.
\item[(i)] Let $T_n$ be the total amount of medicine produced in a week. Find the mean and variance of $T_n$. 
\item<2->[] {\color{red} $Mean(T_n) =   5\times 500 = 2500$}

 {\color{red} $Var(T_n) = 5\times 900 = 4500$}
\item<3->[(ii)] Let $\bar{X}$ be the average amount of medicine produced in a 5-day week. Find the mean and variance of $\bar{X}$.
\item<4->[] {\color{red} $Mean(\bar{X}) = 500$} 

{\color{red} $Var(\bar{X}) = 900/5 = 180$}
\end{itemize}
\end{frame}


\begin{frame}{Questions}
\begin{itemize}
\setlength{\itemsep}{10pt}
\item<1->[Q6:] Suppose the company producing a medicine has different means and variances the amount produced on each day of the week:
\begin{table}
\begin{tabular}{l|ll}
Day & Mean & Variance \\ \hline
Monday $(X_1)$ & 450 & 1200 \\
Tuesday $(X_2)$ & 550 & 800 \\
Wednesday $(X_3)$ & 600 & 500  \\
Thursday $(X_4)$ & 550 & 800 \\
Friday $(X_5)$ & 350 & 1200 
\end{tabular}
\end{table}
\item<2->[] Find mean and variance of both the sum $T_n$ and the average $\bar{X}$.
\item<3->[A6:] {\color{red} $X_1, X_2, X_3, X_4$, and $X_5$ are no longer $i.i.d.$!\\
\bigskip
$Mean(T_n) = 450 + 550 + 600 + 550 + 350 = 2500$

$Var(T_n) = 1200 + 800 + 500 + 800 + 1200 = 4500$}
\item<4->[] {\color{red} $Mean(\bar{X}) = 1/n \times Mean(T_n) = 500$ 

$Var(\bar{X}) = 1/n^2 \times Var(T_n) = 4500/25 = 180$}
\end{itemize}
\end{frame}

\begin{frame}{Questions}
\begin{itemize}
\setlength{\itemsep}{15pt}
\item<1->[Q7:] Let $P_1$ be the proportion of heads in 50 coin tosses, where $P(H) = 0.6$. Find $Mean(P_1)$ and $Var(P_1)$.
\item<2->[A7:] {\color{red} $Mean(P_1) = 0.6$ and $Var(P_1) = 0.6 \times 0.4/ 50 = 0.0048$.}
\item<3->[Q8:] Let $P_2$ be the proportion of heads in 20 coin tosses, where $P(H) = 0.7$. From earlier, $Mean(P_2) = 0.7$ and $Var(P_2) = 0.0105$. Let $D = P_1 - P_2$. Find the mean and variance of $D$.
\item<4->[A8:] {\color{red} $Mean(D) = 0.6 - 0.7 = -0.1$

$Var(D) = 0.0048 + 0.0105 = 0.0153$}
\end{itemize}
\end{frame}



\end{document}


